---
title: "Assign DFO Regions to WSDB records"
author: "LJ Feyrer, J. Stanistreet"
date: "Updated: `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

This R Markdown document is made to interactively update and view the DFO region where sightings records occur prior to import in the WSDB using leaflet 

### REQUIREMENTS:

1. An input file - formatted Excel sheet with two tabs:
    (A) a sighting data sheet prepped with clean coordinate (LATITUDE/LONGITUDE) and species code (SPECIES_CD) info that will be used for DFO region assignment and will be checked for "land whales."
    (B) a "species codes" tab in the input sheet or separate xlsx in the input folder with a tab called "species codes." This defines the list of species codes and associated species names  (This is currently specified as a separate xlsx in "species" object below)

The name of input files needs to be manually updated in this Rmarkdown script in the code block below --->

```{r input file, echo=T}

# INPUTS NEED TO BE CSV files, file names are specified in quotations here:
input_file <-"CoriolisMultiBeamValidatedEntry2023.csv"
species = "speciesCodes.csv" 

```
2. The correct folder structure and file names:
  - *input*: folder contains Excel sheet of sighting coordinates to check land and assign to regions. The name of this file needs to be specified by editing the rmarkdown file in the code block above. 

  - *output*: folder containing .csv files of results: 
    - *FileName-REGIONCODES.csv* - The data with regions, land whales and species code variables populated
    - *FileName-possible_Dups.csv* - The records that may be possible duplicates based on matching coordinates and Date/Time stamps
    - *FileName-NA_COORDS.csv* - The records that have NA values or errors with given coordinate info   
    - *FileName-MULTIPLE_REGIONS.csv* - Any records that fall on the line between regions and have two region assignments (rare, maybe impossible with current analysis, but just in case!)
    
If you want an html map file and subfolder of html files for use outside of R, change save_map to TRUE
  
```{r save html map, echo=T}
save_map = F
```

  - *shapefiles*: folder containing the following shapefiles:
  
      - DFO NAFO EEZ Regions: these are based on NAFO regions (2022), with Gulf/ QC 
        input and manually edited to remove overlap with land, extend to edge of Canada's EEZ, snap edges together, etc.
      - Canada: land area shapefile, sourced from National Atlas files on opengov.ca

```{r assign source region to output csv, echo=FALSE}
source("scripts/assignRegions_pt1.R")
```

### Confirm DFO REGION of the points on a map

```{r visualize region on map, echo=FALSE}
source("scripts/assignRegions_pt2.R")
WSmap1
# 
# saveWidget(WSmap1, file="WSDB_Regions.html")

```

### Region boundary spatial data sources:

|     MAR - From DFO Open Gov region shapefile + NAFO 
|     QC - From DFO Open Gov region shapefile + input from GULF 
|     NL - From DFO Open Gov region shapefile + NAFO 
|     GULF - From DFO Open Gov region shapefile + input from GULF 
|     AR - From DFO Open Gov region shapefile + NAFO 0A/0B
|     O&P - From DFO Open Gov land region shapefile
|     PAC - From DFO Open Gov land region shapefile & EEZ
|     OTHER - Other - anywhere outside Canada EEZ (e.g., St. Pierre)


### Visual check for errors if any of the points are on land. 
##### This will also be noted in the csv output in the variable field "land" for any records that overlap with the land polygon shapefile.

```{r visualize land whales on map, echo=FALSE}
WSmap2
```

